### 2024-02-17 skripsi/abstract.tex
Layout & deriv deeponetsvm

%\chapter{ABSTRACT}
\clearpage
\phantomsection
\addcontentsline{toc}{chapter}{ABSTRACT}
\begin{center}
\textbf{\large  ABSTRACT}\\[0.5cm]
\textbf{\large TITLE}\\[0.5cm]
\textbf{by}\\
\textbf{\penulis}\\
\textbf{NIM: \nim}\\[2em]
\end{center}

\noindent In general, Abstract is a translation of Abstrak. However, appropriate paraphrase may need some words or sentences whose meanings are close enough to those written in Abstrak.

\noindent Key words: \keywords