### 2025-01-09 skripsi/chapter5.tex

% !TEX root = ./skripsi.tex
\chapter{CONCLUSIONS AND FUTURE WORK}

\section{Conclusion}
\noindent Based on the results of the research into \lower{\judul}, several conclusions can be drawn.
\begin{enumerate}
    \item The proposed SpectralSVR model is able to learn the relationships defined by PDEs. This is done with a computational model that with four phases which are data preparation, data transformation, model training, and finally model evaluation. 
    \item Interpretation of the proposed model can be done to a certain extent using two tools which are the correlation image and p-matrix developed by \textcite{ustunVisualisationInterpretationSupport2007}.
    \item The proposed model is also capable in dealing with noise and partial data as shown in scenario 1 and 3.
    \item The model's capabilities are also verified using exact solution. This revealed that for some configurations the model is able to successfully generalize to exact solutions.
\end{enumerate}
% \noindent Bab ini merupakan pamungkas berupa rincian rangkuman yang merupakan simpulan dari analisis yang telah dilakukan. Simpulan ini menyajikan sejumlah hal penting yang disampaikan secara ringkas, padat, dan utuh, yang menjawab tujuan penelitian yang dituliskan pada Bab Pendahuluan. Sangat mungkin   ada beberapa konsekuensi dan implikasi yang ditimbulkan dari simpulan yang dihasilkan, yang sepatutnya menjadi perhatian pada penelitian berikutnya. Judul bab dapat disesuaikan, namun umumnya ada \textit{Simpulan} yang memang mendominasi isi bab ini.

\section{Future Work}
During the work on this study, some limitations were imposed, and many new questions also went unanswered. Here, we would like to list some potential avenues for future work:
\begin{itemize}
    % TODO: reword these points
    \item Most of the synthetic data used to in scenario 1 and 2 were not realistic as they were generated using MMS. The addition of more realistic data generated by traditional numerical methods or other sources could improve the model's performance and .
    \item Test how auto regression of weather is compared to the target data
    \item Compare inverse prediction results with traditional derivatives such as the spectral method or FDM
    \item Compare the proposed method to other methods in comparable situations
    \item Measure and compare the efficiency (time) of computing solutions compared to traditional and machine learning-based methods
    \item Add more basis functions like spherical harmonics or wavelet basis
    \item Add more regression models, specifically more support vector regression models to continue being able to use the correlation image and p-matrix
    \item Use better metrics for measuring performance such as the \enquote{standard error of regression}
    \item Analyze the performance with different hyperparameters or data sizes and plot the result in charts.
    \item Analysis of the error by wave number.
    \item Solve the problem of double penalty that causes the model to tend towards smooth solutions \autocite{brownForecastsSpatialFields2011,NIPS2017_44a2e080}. Try to use more regularization (higher regularization parameter) for higher wave numbers.
    \item Add support for sparse data by using data assimilation techniques like the Lomb-Scargale Periodogram \autocite{vanderplasUnderstandingLombScargle2018}, and 4d-var data assimilation \autocite{puNumericalWeatherPrediction2018,parkDataAssimilationAtmospheric2013}.
    \item A study of the contribution patterns for different terms in a PDE shown by the p-matrix or other methods.
\end{itemize}
% \noindent Sejumlah ide yang muncul ketika melaksanakan penelitian TA dapat menjadi bahan atau topik untuk pekerjaan selanjutnya. Hal ini dapat berupa perbaikan atau ragam lain dari apa yang telah dilakukan sepanjang penelitian. Sub bab ini menjadi sumber informasi penting bagi, utamanya mahasiswa, yang akan melakukan penelitian lanjutan.