### 2025-01-13 skripsi/abstract.tex
fix fourier derivatives with periods, fix typos in cover, change judul to lower case and add title case variant, add toc, tof, tot, toa to table of contents, fix font issue with mathptmx, add statement of originality, remove list of notation (for now), fix incorrect statement about burgers eq in ch2, fix sattelite coverage image caption too long, move burgers eq exact solution to ch2, fix figure positioning H not h, reword future work

% !TEX root = ./skripsi.tex
\clearpage
\phantomsection
\addcontentsline{toc}{chapter}{ABSTRACT}
\begin{center}
\textbf{\large  ABSTRACT}\\[0.5cm]
\textbf{\large \judulcapitalized}\\[0.5cm]
\textbf{by}\\
\textbf{\penulis}\\
\textbf{NIM:\@\nim}\\[2em]
\end{center}

\noindent Numerical models of systems are a crucial part of science and engineering. The use of machine learning in this space for operator learning provides an alternative as data-driven surrogates. The Fourier Transform provides a key component for learning the relationship between a function and its derivatives. Building on Spectral Neural Operators (SNO), we propose a Support Vector Machine (SVM) based framework to learn the underlying governing equations of a system based on data. We study the viability and interpretability of the proposed framework on the derivative equation and the Burgers' equation. The model is able to learn from mathematically correct random data and is able to partially generalize to an exact solution of the Burgers' equation. The learned model is interpreted and verified to have learned the correct contributions of the input function coefficients to the output function coefficients.

\noindent Keywords: \keywords{}