### 2025-01-13 skripsi/statementoforiginality.tex

% !TEX root = ./skripsi.tex
\clearpage
\phantomsection{}
\addcontentsline{toc}{chapter}{STATEMENT OF ORIGINALITY}
\begin{center}
 \textbf{\large STATEMENT OF ORIGINALITY}\\[3em]
\end{center}
%-----------------------------------------
\noindent Hereby, I declare that the thesis titled  
\enquote{\judulcapitalized} and all its contents are entirely my own work. I have not engaged in plagiarism or citation practices that contravene the ethical standards of academic conduct upheld within the scholarly community. By making this declaration, I accept full responsibility and am prepared to face any consequences or sanctions should any violation of academic ethics or claims regarding the originality of this work arise in the future.

\begin{flushright}
Bandung, \today{}\\[2cm]
\penulis{}\\
\nim{}
\end{flushright}