### 2025-01-14 technical_doc/requirements.tex

% !TEX root = ./technical_doc.tex
\chapter{REQUIREMENTS}