### 2025-01-14 technical_doc/usecase.tex

% !TEX root = ./technical_doc.tex
\chapter{USE CASE}