### 2025-02-10 skripsi/identitas.tex
update title to using instead of with, add more to acknowledgements

% !TEX root = ./skripsi.tex
%isi identitas diri di file ini......
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%JUDUL: 
\providecommand{\judul}{approximating solutions of governing equations using Fourier transform and support vector machine}
\providecommand{\judulcapitalized}{Approximating Solutions of Governing Equations Using Fourier Transform and Support Vector Machine}
%PENULIS:
\providecommand{\penulis}{Ahmad Izzuddin}
%NIM:
\providecommand{\nim}{1908919}

%Pembimbing1 dan pembimbing2, misalnya:
%Pembimbing 1
\providecommand{\pembimbingsatu}{Prof.\ Dr.\ Lala Septem Riza, M.T.}
%NIP Pembimbing 1:
\providecommand{\NIPpbbsatu}{197809262008121001}
%Pembimbing2
\providecommand{\pembimbingdua}{Dr.\ Muhammad Nursalman, M.T.}
%NIP Pembimbing 2:
\providecommand{\NIPpbbdua}{197909292006041002}

%untuk keywords yang ada di abstrak, di sini:
\providecommand{\katakunci}{\textit{Operator Learning}, PDE, LSSVM.}
\providecommand{\keywords}{Operator Learning, PDE, LSSVM.}